The Implementation of China’s Overseas NGO Law and the Operating Space for International Civil Society

### 2023-11-21 manuscript/manuscript.qmd
Merge my stuff with Meng's stuff

In the past decade, nearly 100 countries have proposed or enacted legal measures designed to restrict civil society [@ICNL:2021], and only 21% of countries have open and unrestricted civic space [@ChaudhryHeiss:2022a; @Civicus:2023]. This global trend of closing or shrinking civic space [@Chaudhry:2022; @ChaudhryHeiss:2021; @CarothersBrechenmacher:2014; @DupuyPrakash:2018; @DupuyPrakash:2020; @DupuyRonPrakash:2016] has affected countries of all regime types, including both autocracies and democracies such as Morocco and Palestine [@AtiaHerrold:2018], India [@press-trust-india:2019], Vietnam [@Sidel:2023], Canada [@CIVICUS:2013], and the United States [@Cramer:2020]. On April 28, 2016, China passed its own set of legal restrictions on civil society with its Law of the People's Republic of China on Administration of Activities of Overseas Nongovernmental Organizations in the Mainland of China, or "ONGO Law," which formally took effect on January 1, 2017. The laws requires all international NGOs (INGOs) to either register representative offices (ROs) or file temporary activities to lawfully operate in mainland China.As an authoritarian state with a growing nonprofit sector, the Chinese model of implementing restrictive laws for international NGOs (INGOs) has important implications for understanding the dynamics of how civil society laws can be used as policy tools to harness INGOs.

@@

@@

```{r fig-ro-ta-pathways, fig.width=6, fig.height=3}
#| echo: false
#| include: true
#| warning: false
#| message: false
#| fig-cap: "Two possible pathways for international NGOs to legally work in 
#|   China, either as formal Registered Organizations (ROs) or by undertaking 
#|   specific Temporary Activities (TAs)"
ro_nodes <- tibble(
  step = c(
    "Find a professional supervisory unit (PSU)",
    "Obtain PSU approval",
    "Open organization bank accounts",
    "Keep programming withing approved scope",
    "Submit annual work plans (×2)",
    "Submit annual reports (×2)"
  ),
  type = "Registered organization (RO)"
)
ro_edges <- tibble(from = 1:5, to = 2:6)
ro_graph <- tbl_graph(nodes = ro_nodes, edges = ro_edges)

ta_nodes <- tibble(
  step = c(
    "Find an eligible Chinese partner to file paperwork",
    "Obtain PSU approval",
    "Channel money through partner’s bank",
    "Keep programming withing approved scope",
    "Finish work within a year",
    "Submit final project report"
  ),
  type = "Temporary activities (TA)"
)
ta_edges <- tibble(from = 1:5, to = 2:6)
ta_graph <- tbl_graph(nodes = ta_nodes, edges = ta_edges)

@@

::: callout-warning
Apologies! This section is still *incredibly* drafty!
:::

@@

## China's INGO as a policy tool 
*note* changed the sub-title to incorporate the policy tool lense too.

@@

```{r fig-theory, fig.width=6, fig.height=2}
#| echo: false
#| include: true
#| warning: false
#| message: false
#| fig-cap: "Diagram of our theoretical framework showing three possible determinants of INGO operational flexibility: the contentiousness of the INGO's issue area, the inherent geographic scope of the INGO's operations, and the government's interest in and relationship with the INGO."
theory_nodes <- tribble(
  ~id, ~name, ~x, ~y, ~highlight,
  1,   "Issue contentiousness", 1, 0, FALSE,
  2,   "Operational flexibility", 3, 0, TRUE,
  3,   "Local focus", 2, 1, FALSE,
  4,   "Government relationship and interest", 2, -1, FALSE
)

theory_edges <- tribble(
  ~from, ~to,
  1, 2,
  3, 1,
  3, 2,
  1, 4,
  4, 2
)

theory_graph <- tbl_graph(nodes = theory_nodes, edges = theory_edges)

ggraph(theory_graph, x = x, y = y) + 
  geom_edge_link(
    aes(
      end_cap = label_rect(node2.name, padding = margin(10, 5, 10, 5, "pt"))
    ),
    arrow = arrow(angle = 30, length = unit(6, "pt"), type = "closed"),
    edge_linewidth = 0.25
  ) +
  geom_node_label(
    aes(x = x, y = y, label = name, fill = highlight), 
    size = 3, label.r = unit(0, units = "pt"), fontface = "bold"
  ) +
  scale_x_continuous(expand = expansion(add = 0.3)) +
  scale_y_continuous(expand = expansion(add = 0.2)) +
  scale_fill_manual(values = c("white", clrs[1]), guide = "none") +
  labs(x = NULL, y = NULL) +
  theme_ongo() +
  theme(
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    axis.text = element_blank(),
  )
```

@@

-   **Expectation 1** (issue area): Because they are seen as less of a threat to regime stability, INGOs working on less contentious issues should be registered in more provinces

@@

-   **Expectation 2** (local connections): Because they are explicitly focused on local issues, INGos with more local connections should be registered in fewer provinces

Finally, we also expect that the operational space for INGOs is negatively associated with the time elapsed since the law took effect:

-   **Expectation 3** (registration timing): Because formal registration can take a long time and can involve many bureaucratic hurdles, INGOs that registered soon after the law took effect should be be (1) less of a threat and (2) a higher priority for processing, and so should be registered in more provinces

# Data and methods

@@

```{r fig-registration-timeline, fig.width=6, fig.height=3}
#| include: true
#| echo: false
#| fig-cap: "Count of international NGO registrations until the 2017 ONGO Law, divided by quarters"
ggplot(ongo, aes(x = registration_date)) +
  geom_histogram(binwidth = 365.25/4, boundary = ymd("2017-01-01"), 
                 fill = clrs[6], color = "white") +
  scale_x_date(date_breaks = "years", date_labels = "%Y") +
  labs(x = NULL, y = "Count of new INGO registrations") +
  theme_ongo()
```

>>>>

In the past decade, nearly 100 countries have proposed or enacted legal measures designed to restrict civil society [@ICNL:2021], and only 21% of countries have open and unrestricted civic space [@ChaudhryHeiss:2022a; @Civicus:2023]. This global trend of closing or shrinking civic space [@Chaudhry:2022; @ChaudhryHeiss:2021; @CarothersBrechenmacher:2014; @DupuyPrakash:2018; @DupuyPrakash:2020; @DupuyRonPrakash:2016] has affected countries of all regime types, including both autocracies and democracies such as Morocco and Palestine [@AtiaHerrold:2018], India [@press-trust-india:2019], Vietnam [@Sidel:2023], Canada [@CIVICUS:2013], and the United States [@Cramer:2020]. On April 28, 2016, China passed its own set of legal restrictions on civil society with its Law of the People's Republic of China on Administration of Activities of Overseas Nongovernmental Organizations in the Mainland of China, or "ONGO Law," which formally took effect on January 1, 2017. The law requires all international NGOs (INGOs) to either register representative offices (ROs) or file temporary activities to lawfully operate in mainland China.As an authoritarian state with a growing nonprofit sector, the Chinese model of implementing restrictive laws for international NGOs (INGOs) has important implications for understanding the dynamics of how civil society laws can be used as policy tools to harness INGOs.

@@

In this paper, we first explore the ongoing phenomena of closing civic space and authoritarian regulation of civil society, examining how governments conceptualize the role of international civil society in their domestic politics and why governments select regulatory regimes for international NGOs. We then investigate how international NGOs have responded and adapted to more restrictive regulatory environments and theorize that a key component of INGO regulatory adaptation is *organizational flexibility*, or the ability to operate with greater latitude in host countries. We propose two key determinants of this flexibility: (1) organizational characteristics and (2) host government preferences and relationships TODO and their interplay/interaction. We illustrate this theory using the case of China TODO stats and case studies, mixed method, demonstrating how INGOs that address issues that align with government preferences enjoy the most organizational flexibility in the country.

# Closing civic space

## The global constriction of civic space

Many different reasons for these restrictions and several purposes - INGOs pose a threat to regime stability, but they also provide governments with services and technical expertise. Civil society/CSO/NGO + host government definition - De Mattee and institutional taxonomy. My stuff on gatekeeping vs. program capture.

INGO regulations are implemented and enforced differentially - difference in de jure and de facto. This differential enforcement is one mechanism to allow host governments to harness the aspects of INGO programming that align with their domestic policy preferences while limiting INGO behavior that could threaten regime stability.

Why pass laws in the first place and why enforcement differently?

"We discuss the role of civil society organizations (CSOs) as agents of democratization and note the emergence of dual, at times apparently conflicting policy postures within authoritarian regimes (restriction and repression for some CSOs vs. financial support and opportunities for collaboration for others)." [@ToeplerZimmerFrohlich:2020]

Also, [@Plantan:2022] finds "that I find" evidence of selective implementation that reveals which groups are seen as threatening or beneficial."

The presence of de jure regulations and the strictness of their de facto enforcement are both determined by the interplay between INGO characteristics and government preferences.

## Closing civic space in China

China provides a fascinating example of both authoritarian preferences for INGO regulations and INGO adaptations to these regulations.

@@

```{r fig-ro-ta-pathways, fig.width=6, fig.height=3}
#| echo: false
#| include: true
#| warning: false
#| message: false
#| fig-cap: "Two possible pathways for international NGOs to legally work in 
#|   China under the 2017 ONGO law, either as formal Registered Organizations 
#|   (ROs) or by undertaking specific Temporary Activities (TAs)"
ro_nodes <- tibble(
  step = c(
    "Find a professional supervisory unit (PSU)",
    "Obtain PSU approval",
    "Open organization bank accounts",
    "Keep programming within approved scope",
    "Submit annual work plans to PSU and Public Security Bureau",
    "Submit annual reports to PSU and Public Security Bureau"
  ),
  type = "Registered organization (RO)"
)
ro_edges <- tibble(from = 1:5, to = 2:6)
ro_graph <- tbl_graph(nodes = ro_nodes, edges = ro_edges)

ta_nodes <- tibble(
  step = c(
    "Find an eligible Chinese partner to file paperwork",
    "Obtain PSU approval",
    "Channel money through partner’s bank",
    "Keep programming within approved scope",
    "Finish work within a year",
    "Submit final project report to Public Security Bureau"
  ),
  type = "Temporary activities (TA)"
)
ta_edges <- tibble(from = 1:5, to = 2:6)
ta_graph <- tbl_graph(nodes = ta_nodes, edges = ta_edges)

@@

@@

## China's INGO as a policy tool

*note* changed the sub-title to incorporate the policy tool lense too.

@@

```{r fig-theory, fig.width=6, fig.height=2}
#| echo: false
#| include: true
#| warning: false
#| message: false
#| fig-cap: "Diagram of our theoretical framework showing three possible determinants of INGO operational flexibility: the contentiousness of the INGO's issue area, the inherent geographic scope of the INGO's operations, and the government's interest in and relationship with the INGO."
theory_nodes <- tribble(
  ~id, ~name, ~x, ~y, ~highlight,
  1,   "Issue contentiousness", 1, 0, FALSE,
  2,   "Operational flexibility", 3, 0, TRUE,
  3,   "Local focus", 2, 1, FALSE,
  4,   "Government preferences", 2, -1, FALSE
)

theory_edges <- tribble(
  ~from, ~to,
  1, 2,
  3, 1,
  3, 2,
  1, 4,
  4, 1,
  4, 2
)

theory_graph <- tbl_graph(nodes = theory_nodes, edges = theory_edges, directed = TRUE)

ggraph(theory_graph, x = x, y = y) + 
  geom_edge_link(
    aes(
      start_cap = label_rect(node1.name, padding = margin(10, 5, 10, 5, "pt")),
      end_cap = label_rect(node2.name, padding = margin(10, 5, 10, 5, "pt"))
    ),
    arrow = arrow(angle = 30, length = unit(6, "pt"), type = "closed"),
    edge_linewidth = 0.25
  ) +
  geom_node_label(
    aes(x = x, y = y, label = name, fill = highlight), 
    size = 3, label.r = unit(0, units = "pt"), fontface = "bold"
  ) +
  scale_x_continuous(expand = expansion(add = 0.3)) +
  scale_y_continuous(expand = expansion(add = 0.2)) +
  scale_fill_manual(values = c("white", clrs[1]), guide = "none") +
  labs(x = NULL, y = NULL) +
  theme_ongo() +
  theme(
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    axis.text = element_blank(),
  )
```

Based on this framework, we can identify and test several theoretical expectations:

@@

In 2018 Batke and Hang posited that the ONGO law was showing signs of reorienting the missions of legally recognized INGOs and the overall scope of international civil society to be more explicitly aligned with government preferences. @HolbigLang:2022 recently confirmed this prediction, finding that the issue areas addressed by legally recognized INGOs have "shift[ed] towards fields of activity high up the government’s domestic policy agenda" [-@HolbigLang:2022, p. 587].

Following our theory, government preferences thus directly influence INGO issue areas, which in turn shapes the amount of flexibility enjoyed by the INGO. Accordingly, we expect that INGOs working on non-contentious, regime-aligned issues will have greater reach throughout the country:

- **Theoretical expectation 1**: Because they are seen as less of a threat to regime stability and more aligned with government policy preferences, INGOs working on less contentious issues should be registered in more provinces.

@@

- **Theoretical expectation 2**: Because they are explicitly focused on local issues, INGOs with explicitly local missions should be registered in fewer provinces.

Finally, we also expect that the operational space for INGOs is negatively associated with the time elapsed since the law took effect. The mechanism behind this expectation is directly related to both issue area and government preferences.

- **Theoretical expectation 3** (registration timing): Because formal registration can take a long time and can involve many bureaucratic hurdles, INGOs that registered soon after the law took effect should be be (1) less of a threat and (2) a higher priority for processing, and so should be registered in more provinces.

# Data and empirical approach

We test these theoretical expectations using a mixed methods approach. Using

@@

```{r fig-registration-timeline, fig.width=6, fig.height=3}
#| include: true
#| echo: false
#| fig-cap: "Count of international NGO registrations under the 2017 ONGO Law, divided by quarters"
ggplot(ongo, aes(x = registration_date)) +
  geom_histogram(binwidth = 365.25/4, boundary = ymd("2017-01-01"), 
                 fill = clrs[6], color = "white") +
  scale_x_date(date_breaks = "years", date_labels = "%Y") +
  labs(x = NULL, y = "Count of new INGO registrations") +
  theme_ongo()
```
